---
title: "R Notebook"
output: html_notebook
---

# Probability Theory 

Generate 10 separate random flips with probability 0.3
```{r}
rbinom(10, 1, 0.3)

```

Generate 100 occurrences of flipping 10 coins, each with 30% probability
```{r}
rbinom(100,10,0.3)
```

If you have 10 coin flips and there's a 0.5 probability of getting a success 
and a success earns you $1 and a fail costs you \$1, do you take the bet?

X ~ Binom(10, 0.5)
Pr(X = 5)

```{r}
flips <- rbinom(1000000, 10, 0.5)
hist(flips)
pr_5 <- mean(flips == 5)
```

With dbinom you can combine this into one step. What is the probability of
5 heads? 6 heads? 10 heads?

```{r}
pr_5 <- dbinom(5,10, 0.5) # condition is 5, 10 draws, 0.5 prob of success
pr_5
pr_6 <-dbinom(6,10,0.5)
pr_6
pr_10 <- dbinom (10, 10, 0.5)
pr_10
```


Calculate the number of flips <= 4 with 10 flips and a 0.5 probability of success?
X ~ Binom (10, 0.5)
Pr (X <= 4)
```{r}
flips <- rbinom(1000000, 10, 0.5)
cumulative <- mean (flips <= 4)
cumulative

cumulative <- pbinom (4, 10, 0.5)
cumulative
```


If you flip 10 coins each with a 30% probability of coming up heads, what is 
the probability exactly 2 of them are heads?
```{r}
# X ~ Binomial (10, 0.3)
# Pr (X = 2)

# With dbinom
P <- dbinom (2, 10, 0.3)
P

# Confirmation
flips <- rbinom (100000, 10, 0.3)
P <- mean (flips  == 2)
P

```


Calculate the probability that at least five coins are heads
Confirm your answer with a simulation of 10,000 trials
```{r}
p <- 1-pbinom (4, 10, 0.3)
p

flips <- rbinom (10000, 10, 0.3)
p <- mean(flips >= 5)
p
```


What is the expected value of a binomial distribution where 25 coins are 
flipped, each having a 30% chance of heads?
```{r}
# Calcualte the expected value using the exact formula
# X ~ Binom (25, 0.3)
# E[X] = 25 * (0.3)
# Var (X) = 25 * 0.3 * 0.7

e <- 25*0.3
e

#Confirm with a simulation using R binom
e <- mean(rbinom(10000, 25, .3))
e

# Calculate the variance using the exact formula
v <- 25 * 0.3 * 0.7
v

# Confirm with a simulation using rbinom
flips <- rbinom(10000, 25, 0.3)
v <- var(flips)
v
```

# Simulate two coins being flipped 

What is the probability of both A and B being successful?

```{r}
A <- rbinom(10000, 1, 0.5)
B <- rbinom(10000, 1, 0.5)

mean(A & B)
```


If event A has 10% probability and B has a probability of 70% success, what is 
the probability of them occurring together?

```{r}
A <- rbinom (10000, 1, 0.1)
B <- rbinom (10000, 1, 0.7)

mean(A & B)
```


Coin A has a 40% probability of being heads and Coin B has a 20% probability of
being heads, what is the probability of both coming up heads?

```{r}
# Simulate 100,000 flips of a coin with a 40% chance of heads
A <- rbinom(10000, 1, 0.2) 

# Simulate 100,000 flips of a coin with a 20% chance of heads
B <- rbinom (10000, 1, 0.4)

# Estimate the probability both A and B are heads
mean(A & B)
```


What if a 3rd coin, C, is tossed with a 70% chance of heads? What is the 
probability of all 3 occurring? 

```{r}
# Theoretically
P <- 0.2*0.4*0.7
P

C <- rbinom(100000, 1, 0.7)
mean (A & B & C)
```

What is the probability of either of two coins being flipped to come up heads 
if both have a 50% probability of coming up heads?
```{r}

# Theoretically,
# P(A | B) = 0.5 + 0.5 - 0.5 * 0.5 = 7.5%

A <- rbinom(100000, 1, 0.5)
B <- rbinom(100000, 1, 0.5)
mean (A | B)
```


What if A has a 20% prob and B has a 40% prob?
```{r}
# Theoretically
# P(A | B) = 0.2 + 0.4 - 0.2 * 0.4 = 0.6 - 0.08 = 0.52

A <- rbinom (100000, 1, 0.2)
B <- rbinom (100000, 1, 0.4)
mean(A | B)
```

If coins A and B are independent, and A has a 60% chance of coming up heads, 
and event B has a 10% chance of coming up heads, what is the probability 
either A or B will come up heads?
```{r}
A <- rbinom(100000, 1, 0.6)
B <- rbinom(100000, 1, 0.1)
mean(A|B)
```


Suppose X is a random Binom(10, .6) variable (10 flips of a coin with 60% 
chance of heads) and Y is a random Binom(10, .7) variable (10 flips of a coin 
with a 70% chance of heads), and they are independent.

What is the probability that either of the variables is less than or equal to 4?
```{r}
X <- rbinom (100000, 10, 0.6)
Y <- rbinom (100000, 10, 0.7)

X_mean = mean(X <= 4)
X_mean
Y_mean = mean(Y <= 4)
Y_mean

P <- X_mean + Y_mean - X_mean * Y_mean
P

P <- mean (X <= 4 | Y <= 4)
P
```

If X is a binomial with size 50 and p = .4, what is the expected value of 3*X?
```{r}
# Theoretically
# E[k*X] = constant * sample_size * probability of success
p <- 3 * 50 * 0.4
p

# Programmatically
X <- rbinom (100000, 50, 0.4)
mean(X* 3) 

```

Simulate 100,000 draws of X, a binomial random variable with size 20 and p = .1. Save this as X
Use this simulation to estimate the expected value of X.
Use this simulation to estimate the expected value of 5*X, as well.
Use this simulation to estimate the variance of X.
Estimate the variance of 5 * X
```{r}
X <- rbinom(100000, 20, 0.1)
pX<- mean(X)
p5X <- mean(X) * 5
pX
p5X

vX <- 20 * 0.1 * 0.9
v5X <- 20 * 0.1 * 0.9 * 5**2
vX
v5X

var(X)
var(X * 5)

```

Calculate Z ~ X + Y where X is 10 coin flips with a 50% probability of success
and Y is 100 coin flips with a 20% probability of success. What is the point 
estimate of X, Y, and Z? What is the variance?
```{r}
X <- rbinom (100000, 10, 0.5)
Y <- rbinom (100000, 100, 0.2)

EX <- mean(X)
EX

EY <- mean(Y)
EY

Z <- X + Y
EZ <- mean(X + Y)
EZ

VX <- 10 * 0.5 * 0.5
VX

VY <- 100 * 0.2 * 0.8
VY

VZ <- VY + VX
VZ

var(X + Y)
var

hist (X + Y)


```

If X is drawn from a binomial with size 20 and p = .3, and Y from size 40 
and p = .1, what is the expected value (mean) of X + Y?

```{r}
# Theortically 
# E[X] = 20 * 0.3 = 6
# E[Y] = 40 * 0.1 = 4
# Z[Y] = 6 + 4 = 10

X <- rbinom(100000, 20, 0.3)
Y <- rbinom(100000, 40, 0.1)
Z <- mean(X) + mean (Y)
Z

```

Simulate 100,000 draws of X (size 20, p = .3) and Y (size 40, p = .1). Estimate 
the expected value of X + Y. Estimate the variance of X + Y, estimate the
variance of 3*X + Y
```{r}
# Theoretically
# E[X] = 20 * 0.3 = 6
# E[Y] = 40 * 0.1 = 4
# V[X] = 20 * 0.3 * 0.7 = 4.2
# V[Y] = 40 * 0.1 * 0.9 = 3.6
# E[Z] = E[X] + E[Y] = 6 + 4 = 10
# V[Z] = V[X] + V[Y] = 4.2 + 3.6 = 7.8
# V[3X] = 3**2 * V[X] = 37.8
# V[3Z] = V[3X] + V[Y] = 37.8 + 3.6 = 41.4

X <- rbinom (100000, 20,0.3)
Y <- rbinom (100000, 40, 0.1)
Z <- mean(X) + mean(Y)
Z

V <- var(X) + var(Y)
V

V3 <- var(3*X) + var(Y)
V3
```

Given a coin, how do you know if it is loaded? What is the probability if you 
get 16 heads and 4 tails?

```{r}
# Flip the coin 20 times and if the mean probability is not 50% + 1 SD == loaded

# 16 heads
P <- 16/20
P
```


If you have 2 piles of 50,000 coins and you assume  50,000 are fair and 50,000
are biased. if you flipped each coin 20 times, what would the histogram look 
like?

Pr(Biased | 14 Heads) = #biased/w14 heads / #total w/14 heads

```{r}
X <- rbinom(50000,20, 0.5)
hist (X)

Y <- rbinom(50000,20, 0.75)
hist(Y)

hist(X + Y)

# Run the simulation the 50,0000 times and see how many tosses resulted in 
# 14 successes

X_14 <- length(X[X == 14])
X_14
probX_14 <- num_14/50000
probX_14

Y_14 <- length(Y[Y == 14])
Y_14
probY_14 <- Y_14/50000
probY_14

Z_14 <- X_14 + Y_14
Z_14
pZ_14 <- Z_14/50000
pZ_14

p <- Y_14/(X_14 + Y_14)
p
```
We see 11 out of 20 flips from a coin that is either fair (50% chance of heads) 
or biased (75% chance of heads). How likely is it that the coin is fair? 
Answer this by simulating 50,000 fair coins and 50,000 biased coins.

```{r}
f <- rbinom(50000, 20, 0.5)
b <- rbinom(50000, 20, 0.75)
f11 <- length(f[f == 11])
b11 <- length(b[b == 11])
p <- f11/(f11 + b11)
p
```

We see 16 out of 20 flips from a coin that is either fair (50% chance of heads) 
or biased (75% chance of heads). How likely is it that the coin is fair?

```{r}
f <- rbinom(50000, 20, 0.5)
b <- rbinom(50000, 20, 0.75)

f16 <- length(f[f == 16])
b16 <- length(b[b == 16])
p <- f16/(f16 + b16)
p
```


Given a coin, you suspect that there’s a 90% chance that it’s fair. In this 
case there are 90,000 fair coins and 10,000 biased coins. You get 14 heads out 
of 20 trials. Is your coin fair?

```{r}
f <- rbinom (90000, 20, 0.5)
b <- rbinom(10000, 20, 0.75)
f14 <- length(f[f == 14])
b14 <- length(b[b == 14])
f14
b14

fraction_biased = b14/(f14 + b14)
fraction_biased


```


We see 14 out of 20 flips are heads, and start with a 80% chance the coin is 
fair and a 20% chance it is biased to 75%. You'll solve this case with 
simulation, by starting with a "bucket" of 10,000 coins, where 8,000 are fair 
and 2,000 are biased, and flipping each of them 20 times.

```{r}
# Simulate 8000 cases of flipping a fair coin, and 2000 of a biased coin
fair_flips <-rbinom(8000,20,0.5)
biased_flips <-rbinom(2000, 20, 0.75)

# Find the number of cases from each coin that resulted in 14/20
fair_14 <- sum(fair_flips == 14)
biased_14 <-sum(biased_flips == 14)

# Use these to estimate the posterior probability
fair_14/(fair_14 + biased_14)
```


Suppose instead of a coin being either fair or biased, there are three 
possibilities: that the coin is fair (50% heads), low (25% heads), and 
high (75% heads). There is a 80% chance it is fair, a 10% chance it is 
biased low, and a 10% chance it is biased high.You see 14/20 flips are heads. 
What is the probability that the coin is fair?
```{r}
# Simulate 80,000 draws from fair coin, 10,000 from each of high and low coins
flips_fair <- rbinom(80000, 20, 0.5)
flips_high <- rbinom(10000, 20, .75)
flips_low <- rbinom(10000, 20, .25)

# Compute the number of coins that resulted in 14 heads from each of these piles
fair_14 <- sum (flips_fair == 14)
high_14 <- sum(flips_high == 14)
low_14 <- sum(flips_low == 14)

# Compute the posterior probability that the coin was fair
fair_14/(fair_14 + high_14 + low_14)
```











Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Cmd+Option+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Cmd+Shift+K* to preview the HTML file). 

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.

